### 2018-07-05 paper/4-build.tex
Work on S4

This approach requires that a task may be started and then suspended until
another task is complete. This can be done with cheap green threads and blocking
(the original approach of \Shake) or using continuation-passing style (what
\Shake does currently).

@@

Suppose the scheduler decides that that a key should be brought up to date. The
next question is: does any work need to be done, or is it already up to date?
Or, in a cloud-build system, do we have a cached copy of the value we need?
These questions can be addressed in one of four fundamental ways, with a number
of tweaks and variations within them.

@@

With a dirty bit it is simple to achieve minimality. However, to achieve early
cutoff (\S\ref{sec-background-shake}) it would be important to reset the dirty
bit after a computation that did not change the value and make sure that cells
that depend on it are not rebuilt unnecessarily. For \Excel, this is difficult
because the dependent cells have already been recursively marked dirty. For
\Make, it is impossible to mark a file clean and at the same time not mark the
keys that depend on it dirty. \Make can approximate early cutoff by not
modifying the result file, and not marking it clean, but then it will be rebuilt
in every subsequent build.

@@

\noindent
The function \hs{recordCT} looks like \hs{recordVT}, but instead of just passing
the hash of the resulting value, we require the whole value. The \hs{verifyVT}
has been replaced with \hs{constructCT}, which instead of taking the hash of the
current value as \emph{input}, returns a list of suitable values as \emph{output}.
If the current value in the store
matches one of the possible values, the build can skip this rule. If the
resulting list is empty, the key must be rebuilt. However, if the current value
does not match the store, and there is a possible value, we can use any value
from the constructive list \emph{without} doing any work to build it, and copy
it into the store from the cloud.

@@

\item Any \hs{Applicative} build system using constructive traces, e.g.
\Bazel, can index directly from the key and results to the output result~--~i.e.
\hs{Map}~\hs{(@@k,}~\hs{[Hash}~\hs{v])}~\hs{v}. Importantly, assuming the traces
are stored on a central server, the client can compute the key and the hashes of
its dependencies, then make a single call to the server to retrieve the result.

@@

With the information in this section we can build Table~\ref{tab-build-systems},
which tabulates combinations of the scheduling strategy and the out-of-date keys
strategy, providing 12~possible build systems, 8~of which are actually inhabited
by existing build systems (we discuss all these systems in
\S\ref{sec-related-build}). Of the remaining 4~spots, all result in workable
build systems. The most interesting unfilled spot in the table is suspending
constructive traces, which would provide many benefits, and which we title
\Cloud \Shake and further explore further in~\S\ref{sec-implementation-cloud}.
% (as we plan on extending \Shake to occupy that spot)

>>>>

This approach requires that a task may be started and then suspended until
another task is complete. Suspending can be done with cheap green threads and blocking
(the original approach of \Shake) or using continuation-passing style (what
\Shake currently does).

@@

Suppose the scheduler decides that a key should be brought up to date. The
next question is: does any work need to be done, or is it already up to date?
Or, in a cloud-build system, do we have a cached copy of the value we need?
These questions can be addressed in one of four fundamental ways, with a number
of tweaks and variations within them.

@@

With a dirty bit it is possible to achieve minimality. However, to achieve early
cutoff (\S\ref{sec-background-shake}) it would be important to reset the dirty
bit after a computation that did not change the value and make sure that cells
that depend on it are not rebuilt unnecessarily. For \Excel, this is difficult
because the dependent cells have already been recursively marked dirty. For
\Make, it is impossible to mark a file clean and at the same time not mark the
keys that depend on it dirty. \Make can approximate early cutoff by not
modifying the result file, and not marking it clean, but then it will be rebuilt
in every subsequent build.

@@

\noindent
The function \hs{recordCT} looks like \hs{recordVT}, but instead of just passing
the hash of the resulting value, we require the actual value. The \hs{verifyVT}
has been replaced with \hs{constructCT}, which instead of taking the hash of the
current value as \emph{input}, returns a list of suitable values as \emph{output}.
If the current value in the store
matches one of the possible values, the build can skip this rule. If the
resulting list is empty, the key must be rebuilt. However, if the current value
does not match the store, and there is a possible value, we can use any value
from the constructive list \emph{without} doing any work to build it, and copy
it into the store.

@@

\item Any \hs{Applicative} build system using constructive traces, e.g.
\CloudBuild, can index directly from the key and results to the output result~--~i.e.
\hs{Map}~\hs{(@@k,}~\hs{[Hash}~\hs{v])}~\hs{v}. Importantly, assuming the traces
are stored on a central server, the client can compute the key and the hashes of
its dependencies, then make a single call to the server to retrieve the result.

@@

With the information in this section we can build Table~\ref{tab-build-systems},
which tabulates combinations of the scheduling strategy and the out-of-date key
strategy, providing 12~possible build systems, 8~of which are inhabited
by existing build systems (we discuss these systems in~\S\ref{sec-background} and
\S\ref{sec-related-build}). Of the remaining 4~spots, all result in workable
build systems. The most interesting unfilled spot in the table is suspending
constructive traces, which would provide many benefits, and which we title
\Cloud \Shake and explore further in~\S\ref{sec-implementation-cloud}.
% (as we plan on extending \Shake to occupy that spot)
